### 2016-11-23 local-capability-machine.tex
Added clear macros

\subsection*{Semantics}
Assume a $\decode$ function that decodes integer to instructions:
\decode &:\Words \fun \Instrs
Assume an $\encodePerm$, $\encodeLoc$, and $\encodePermPair$ function that encodes a permissions, locality, and permission pair, respectively, as an integer:
\encodePerm &: \Perms \fun \ints \\
\encodeLoc &: \Globals \fun \ints \\
\encodePermPair &: (\Perms \times \Globals) \fun \ints \\
Further, assume an inverse function, $\decodePermPair$, that decodes permissions
\[
  \decodePermPair : \ints \fun (\Perms \times \Globals)
\]
  \Phi & \rightarrow \sem{\decode(\memheap(\addr))}(\Phi) & &                                   
                                                              \begin{array}{l}
                                                                \quad\text{and $\start \leq \addr < \addrend$}\\
                                                                \quad\text{and $\perm \in \{ \exec,\rwx, \rwlx \}$ }
\Phi & \rightarrow \failed                                 & & \text{otherwise}
It is assumed that address 0 is used for I/O, so whatever resides in location 0 after an execution can be seen as a result. As we will see in the semantics, it is assumed that the initial process starts with a readwrite capability for this location.

@@

\begin{description}
\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
  \begin{itemize}
  \item Set up activation record
  \item Create local enter capability for activation
  \item Clear unused registers
  \item Jump
  \item Upon return: Run activation code
  \end{itemize}
A more detailed description of each of the above steps:
\begin{description}
  \item [Set up activation record]
    \begin{itemize}
    \item Run malloc to get a piece of memory with space for:
    \begin{itemize}
      \item Words in $\bar{r}_{\var{priv}}$
      \item Code return capability (opc)
      \item Activation code
    \end{itemize}
    \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
    \item Adjust the a copy of the current pc to point to the return address and save it to the activation record.
    \item Write the activation code to the activation record.
  \end{itemize}
  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability. Place this capability in $r_0$.
  \item [Clear unused registers]
    Clear all the register that are not $\pcreg$, $r_0$ or in $\bar{r}_{\var{args}}$.
  \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
  \item [Activation code] The activation code does the following:
    \begin{itemize}
    \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
%    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
    \item Load the return address to $\pcreg$
    \end{itemize}
\end{description}
\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
\item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$ (if one of them is an integer, then use that in the comparison). If they are equal, then execution continues. If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
\end{description}
Note:
\begin{itemize}
\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local. Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
\end{itemize}
\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.

@@

>>>>

\subsection*{Semantics}
Assume a $\decode$ function that decodes integer to instructions:
\begin{align*}
\decode &:\Words \fun \Instrs
\end{align*}
Assume an $\encodePerm$, $\encodeLoc$, and $\encodePermPair$ function that encodes a permissions, locality, and permission pair, respectively, as an integer:
\begin{align*}
\encodePerm &: \Perms \fun \ints \\
\encodeLoc &: \Globals \fun \ints \\
\encodePermPair &: (\Perms \times \Globals) \fun \ints \\
\end{align*}
Further, assume an inverse function, $\decodePermPair{}$, that decodes permissions
\[
  \decodePermPair{} : \ints \fun (\Perms \times \Globals)
\]
\begin{align*}
  \Phi & \rightarrow \sem{\decode(\memheap(\addr))}(\Phi) & &                                   
                                                              \arraycolsep=0pt
                                                              \begin{array}{l}
                                                                \text{if $\memreg(\pcreg) = \stdcap$}\\
                                                                \quad\text{and $\start \leq \addr < \addrend$}\\
                                                                \quad\text{and $\perm \in \{ \exec,\rwx, \rwlx \}$ }
                                                              \end{array}\\
\Phi & \rightarrow \failed                                 & & \text{otherwise}
\end{align*}

A number of functions and predicates used in the definition of $\sem{-}$ (defined later). Notice all of them are total.

@@

\begin{description}
\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
  \begin{itemize}
  \item Set up activation record
  \item Create local enter capability for activation
  \item Clear unused registers
  \item Jump
  \item Upon return: Run activation code
  \end{itemize}
A more detailed description of each of the above steps:
\begin{description}
  \item [Set up activation record]
    \begin{itemize}
    \item Run malloc to get a piece of memory with space for:
    \begin{itemize}
      \item Words in $\bar{r}_{\var{priv}}$
      \item Code return capability (opc)
      \item Activation code
    \end{itemize}
    \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
    \item Adjust the a copy of the current pc to point to the return address and save it to the activation record.
    \item Write the activation code to the activation record.
  \end{itemize}
  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability. Place this capability in $r_0$.
  \item [Clear unused registers]
    Clear all the register that are not $\pcreg$, $r_0$ or in $\bar{r}_{\var{args}}$.
  \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
  \item [Activation code] The activation code does the following:
    \begin{itemize}
    \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
%    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
    \item Load the return address to $\pcreg$
    \end{itemize}
\end{description}
\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
\item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$ (if one of them is an integer, then use that in the comparison). If they are equal, then execution continues. If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
\item[cclear $r$] Stores 0 to all the memory cells the capability $r$ governs.\footnote{This may in some cases seem like an unreasonable slow instruction. In a real system it would probably be implemented as a vector operation which allows modification of continous segments of memory rather fast.}
\item[rclear $\bar{r}$] Moves 0 to all the registers in the list $\bar{r}$.
\end{description}
Note:
\begin{itemize}
\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local. Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
\end{itemize}
\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.

@@

\item In a real setting due to a limited number of registers, some of the arguments might be spilled to the stack. It would be possible to do something similar here, but to keep
 matters simple, we opt not to do so.
